\papertitle

### 2025-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

The mean of a set of samples is the value that minimises the sum of squared distances between the mean and the samples.~\cite{neilsen_matrix_2012}
The circular weighted average centre of mass satisfies this criterion by minimising the chordal distance between the points on the planar disk formed by the circle.~\cite{hotz_non-asymptotic_2016}
However as outlined above, this position is not physically meaningful and requires projection onto the circular manifold, leading to the inconsistency with the continuum centre of mass.

The position that matches the continuum mean of a periodic system is the intrinsic mean, which minimizes the sum of the squared arc distances between the mean and the samples. 
As all the points exist on the periodic unit circle, all positional information is contained in their angular position. 
This results in a computationally expensive minimisation problem~\cite{hotz_extrinsic_2013}, for which an analytical solution for the intrinsic mean can be found by producing $N$ candidate means,
%
\begin{equation}
\bar{\theta}_{n} = \bar{\phi} + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
\end{equation}
%
where $\bar{\phi}$ represents a naïve angular mass-weighted average position for all particles in the group.
The true intrinsic mean is then the value in the vector, $\bar{\bm{\theta}} = [\bar{\theta}_{1}, \bar{\theta}_{2}, \ldots \bar{\theta}_{n}]$, with the lowest sum of squared arc distances between the points. 
This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_intrinsic_2011} but will always produce the true intrinsic mean.

@@

To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering approach, \cref{fig:method_valid} compares it to the intrinsic weighted average centre of mass. 
As can be seen the pseudo-centre of mass recentering approach produces a perfect estimate of the intrinsic center of mass, provided that particles do not span more half the box. 
Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering approach should be used to ensure accuracy in the centre of mass calculation. 
%
\begin{figure}
    \centering
    \includegraphics[width=8cm]{Fig3.pdf}
    \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighte average centre of mass. 
    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
    Apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the difference in the circular weighted averaging centre of mass causes the point to wrap around to the opposite side of the periodic cell.
    }
    \label{fig:method_valid}
    \script{Fig3.py}
\end{figure}
%

@@

In \cref{fig:sine_wave_assym} we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
This is compared with a system where the particles are symmetrically positioned, in which the circular weighted averaging approach accurately estimates the continuum centre of mass. 
By describing an asymmetric function with a single sine wave, we produce an incorrect estimate of the intrinsic centre of mass. 
First-moment-based methods such as this are known to be sensitive to asymmetry \cite{teague_robust_2018} as the singular sine wave cannot encode asymmetrical information effectively. 
%
\begin{figure}
    \centering
    \includegraphics[width=8cm]{Fig4.pdf}
    \caption{Mass density histograms for two 3 particle molecules, (top) asymmetrically distributed particles [0.26, 0.34, 0.74], (bottom) symmetric molecule [0.26, 0.50, 0.74] (blue bars), and the sine waves produced from the first components of the fourier series (dashed lines). It can be seen in the top image that the asymmetry in the molecule causes the maximum of the sine wave to be a poor estimator of the intrinsic centre of mass. }
    \label{fig:sine_wave_assym}
    \script{Fig4.py}
\end{figure}
%

>>>>

The mean of a set of samples is the value that minimises the sum of the squared distances to the samples.~\cite{neilsen_matrix_2012}
The circular weighted average centre of mass satisfies this criterion by minimising the chordal distance between the points on the planar disk formed by the circle.~\cite{hotz_non-asymptotic_2016}
However as outlined above, this position is not physically meaningful and requires projection onto the circular manifold, leading to the inconsistency with the continuum centre of mass.

The position that matches the continuum mean of a periodic system is the intrinsic mean, which minimizes the sum of the squared arc distances between the mean and the samples. 
As all the points exist on the periodic unit circle, all positional information is contained in their angular position. 
This results in a computationally expensive minimisation problem~\cite{hotz_extrinsic_2013}, for which an analytical solution for the intrinsic mean can be found by producing $N$ candidate means,
%
\begin{equation}
\bar{x}_{n} = \bar{x}_0 + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
\end{equation}
%
where $\bar{x}_0$ represents a naïve mass-weighted average position for all particles in the group.
The true intrinsic mean is then the value in the vector, $\bar{\bm{x}} = [\bar{x}_{1}, \bar{x}_{2}, \ldots \bar{x}_{n}]$, with the lowest sum of squared arc distances between the points. 
This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_intrinsic_2011} but will always produce the true intrinsic mean.

@@

To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering approach, \cref{fig:method_valid} compares it to the intrinsic weighted average centre of mass. 
As can be seen the pseudo-centre of mass recentering approach produces a perfect estimate of the intrinsic center of mass, provided that particles do not span more half the box. 
Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering approach should be used to ensure accuracy in the centre of mass calculation. 
%
\begin{figure}
    \centering
    \includegraphics[width=8cm]{Fig3.pdf}
    \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighted average centre of mass. 
    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
    Apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the difference in the circular weighted averaging centre of mass causes the point to wrap around to the opposite side of the periodic cell.
    }
    \label{fig:method_valid}
    \script{Fig3.py}
\end{figure}
%

@@

In \cref{fig:sine_wave_assym}, we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
This is compared with a system where the particles are symmetrically positioned, in which the circular weighted averaging approach accurately estimates the continuum centre of mass. 
By describing an asymmetric function with a single sine wave, we produce an incorrect estimate of the intrinsic centre of mass. 
First-moment-based methods such as this are known to be sensitive to asymmetry \cite{teague_robust_2018} as the singular sine wave cannot encode asymmetrical information effectively. 
%
\begin{figure}
    \centering
    \includegraphics[width=8cm]{Fig4.pdf}
    \caption{Mass density histograms for two 3 particle molecules, (top) asymmetrically distributed particles [0.26, 0.34, 0.74], (bottom) symmetric molecule [0.26, 0.50, 0.74] (blue bars), and the sine waves produced from the first components of the fourier series (dashed lines). It can be seen in the top image that the asymmetry in the molecule causes the maximum of the sine wave to be a poor estimator of the intrinsic centre of mass. }
    \label{fig:sine_wave_assym}
    \script{Fig4.py}
\end{figure}
%
